The Correctness of Algebraic File Synchronization

### 2015-11-27 2015/paper.tex
Proof for moving inverse sequences

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyx}{\fscommand{Y}{X}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\cxw}{\fscommand{X}{W}}
\newcommand{\czw}{\fscommand{Z}{W}}
\newcommand{\cqr}{\fscommand{Q}{R}}
\newcommand{\cqq}{\fscommand{Q}{Q}}

@@

\newcommand{\eqext}{\sqsubseteq}
\newcommand{\eqnrw}{\sqsupseteq}
\newcommand{\nequiv}{\not\equiv}
\newcommand{\indep}{\,{\wr\wr}\,} % Independent commands, sequences
\newcommand{\unrel}{\indep} % Unrelated nodes
\newcommand{\nindep}{\,{\centernot{\wr\wr}}\,}
\newcommand{\nunrel}{\nindep}

\newcommand{\workssign}{\mathbf{w}}
\newcommand{\works}[1]{\workssign({#1})}
\newcommand{\worksc}[2]{\workssign({#1}|{#2})}

\newcommand{\emptyseq}{[\,]} % TODO change

\newcommand{\extset}[1]{\wp({#1})}
\newcommand{\orderset}[1]{\vec{\wp}({#1})}
\newcommand{\ordered}[1]{#1}
\newcommand{\seqset}[1]{\underline{#1}} % Set of sequences

@@

To model this, in $\setv$ we distinguish between files ($\setf$) and directories ($\setd$), that is,
if $\setb = \{\empt\}$ then:
\[ \setv = \setb \cup \setf \cup \setd \]
The tree propety can then be defined as
\[ \forall n\in\setn : \FS(n) \neq \empt \wedge \parent(n) \neq \topnode \Rightarrow \FS(\parent(n)) \in \setd \]

This means that as we move down from the root of a tree of nodes,
the types of values we encounter in the filesystem can only change according to the following
transition diagram:

@@

% <<<<<<<< TODO Rewrite: no algebra
TODO: Rewrite:
 about which we will reason
using our algebra.
We will aim to draw conclusions or judgements about commands and sequences of commands
based on the algebra we are aiming to construct.
For example, if we believe that two sequences of commands are equivalent as their effect
is the same, then we aim to be able to derive this from the axioms and inference rules of the
algebra.
As the algebra operates on commands only,
these judgements, by definition, need to hold regardless of the filesystem the commands are applied to.
Accordingly, we expect that the more information one encodes into the commands and the sequences,
the better predicitons we will be able to make using our algebra,
as then the commands and sequences will be more specific, and will select a smaller subset
of potential filesystems on which they can meaningfully operate.
% >>>>>>>>>> to here

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(n,V)$, $edit(n,V)$ and $remove(n)$ where $n\in\setn$ and $V\in\setv$ (but $V\neq\empt$).
Regarding what we know of their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC:alg} we know that a useful set of axioms
will in some cases need to distinguish between edits that result in directories ($edit(n,D)$) and
ones that result in files ($edit(n,F)$), and treat them as separate commands.
Indeed, Bill Zissimopoulos suggested (\cite{BZ}) that extending this distinction to more commands
utlimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($D$, $F$ or $\empt$) in the commands is definitely useful.
At the same time, there is never any need to encode information about the
\emph{exact} output value of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

@@

We therefore propose to have a command for each pair of types.
For want of a better system, we will name our commands by concatenating
two of $\fsregcommandchar{D}$, $\fsregcommandchar{F}$ and $\fsregcommandchar{\empt}$, 
where the left sign denotes the type of value
required in the filesystem before the command is applied, and the right sign notes the type
afterwards. For example, $mkdir(n,D)$ will be $\cbd(n,D)$, and $rmdir(n)$ will be $\cdb(n)$.

A command can only succeeds if the original value at node $n$ has a type that matches
the input type of the command. If this is not the case, or if the resulting
filesystem no longer has the tree property, then we say that the command
\textbf{breaks} the filesystem. Broken filesystems are considered equal
(but not equal to any working filesystem), and we note them by $\FS=\fsbroken$.

@@

When we describe multiple or unknown commands, we may substitute one or both
types with a variable, for example by writing $\cxy(n,V)$. In this context,
if we write that $\cxy=\czw$, we mean that their input and output types
are the same ($X=Z$ and $Y=W$).

@@

The commands we aim to investigate do not occur in isolation:
for example, when comparing to filesystems to determine the commands needed to transform one into the other,
we receive a set of commands with uncertain ordering.
Applying commands to a filesystem, however, happens in time,
so for this we need a sequence of commands.
We use $\cc$ to concatenate commands to form a sequence, or concatenate sequences to form a longer sequence:
$S=\cbd(n)\cc \cbf(m,F)$, or $S'=S\cc\cbb(o)$.

@@

Our aim is to derive information about the effects of sequences
of commands independently of that actual filesystems.
In order to make this possible, we investigate the smallest building
blocks of sequences: pairs of commands that act on a filesystem directly one after the other.
This approach is useful as there are a limited number of command pairs,
because, as we arguad above, we can disregard the exact output value of commands apart from its type,
and we can also abstract the relationship between the nodes in the two commands
to a finite number of cases.

@@

Below we list a number of statements derived using this method.
Although they can be proven by investigating all relevant possibilities
in a filesystem, we refer to these in axioms because in a secondary algebraic model
that describes commands without reference to an underlying filesystem,
these would function as meta-rules generating statements we would accept as true.
See section (TODO) for more about such a model.

@@

In this section we describe an algorithm with which all possible orderings of a
set of sequences can be generated, but before doing so, let us consider the following lemma.

\begin{mylem}\label{connected_changes}
If a simple set of commands that can be applied to a filesystem in some order without breaking it
contains commands on $n$ and $\nn$ where $\nn\descendant n$
but $\nn\neq\parent(n)$, then the set must also contain a command on $\parent(n)$.
\end{mylem}
\begin{proof}
We will prove this statement for simple sequences, as if all sequences must contain a command on $\parent(n)$
under the given conditions, then so must sets because otherwise there would be no order in which thay can be
applied to any filesystem.

Let $A$ be a sequence that satisfies the conditions;
we therefore know that it contains a command $\cxy(n,V_Y)$ on $n$
and another command $\czw(\nn,V_W)$ on $\nn$.
We use an indirect proof and assume that there are no commands on $\parent(n)$ in $A$.
Next, we create a new sequence $A'\equiv A$ in which $\cxy$ and $\czw$ are next to each other.
If they are already next to each other in $A$, there is nothing to do.
Otherwise, consider the command next to $\cxy(n,V_Y)$ in the direction where $\czw(\nn,V_W)$ is.
Let this command be $\cqr(m,V_R)$.
If $\czw$ is to the right, the $A$ looks like the following:
\[ A = \cdots\cc\cxy(n,V_Y)\cc\cqr(m,V_R)\cc\cdots\cc\czw(\nn,V_W)\cc\cdots \]
If $n\unrel m$, reverse $\cxy$ and $\cqr$. Based on \cref{ax_separate_commute} we know that the new
sequence is equivalent to $A$.
Otherwise, we know $m\neq\parent(n)$ as there are no commands on $\parent(n)$, and so
from \cref{ax_distantrel_breaks} we get $A\equiv\cbrk$ which contradicts our assumptions.
(Note that $A$ cannot contains assertion commands.)
By repeating this step we can therefore convert $A$ into $A'$ where $\cxy$ and $\czw$ are neighboring commands.
However, then \cref{ax_distantrel_breaks} applies to the sequence and therefore $A\equiv A'\equiv\cbrk$ which
is again a contradiction.
\end{proof}

By applying \cref{connected_changes} multiple times we see that if a simple set of commands
contains two commands on related nodes, then it must contain a command on each
node that is between these two nodes in the parent--child tree.

@@

To order the commands in the component,
select the command with the topmost node, so that there would be no command
on its parent node (if there is one). 
If the command removes content at this node (it is of the form $\fsregcommand{X}{\empt}$), 
then it must follow all other commands in the component, and
all parent--child pairs in the whole component must be destruction pairs.
Otherwise, the command with the topmost node needs to precede the other commands,
and the component must be made of construction pairs.
These follow from the patterns of construction and destruction pairs, which
at no point can mix without causing any sequence to break all filesystems.

Then move the command with the topmost node to the beginning or the end of
the sequence we are creating out of the component,
and recursively start the whole algorithm again on the commands in the component that remain.
The result of this process is a set of possible orders for each component,
which are then combined into a set of possible orders for the whole set of commands.

@@

As a specific case, we can regard a sequence of commands $S$
as a set and determine its $\orderset{S}$.
Let us investigate the properties of simple sequences and this set a bit further.

\begin{mydef}{$\extset{S}$.}
For any sequence $S$ we can consider all sequences that extend it:
$\extset{S} := \{S'|S\eqext S'\}$
\end{mydef}

\begin{mylem}\label{simple_is_max}
For any simple sequence $S\neq\cbrk$, $\forall T: S\eqext T \Rightarrow S\equiv T$;
that is, the domains of simple sequences cannot be extended further.
This is equivalent to the proposition that for simple sequences $\extset{S} = \{S'|S\equiv S'\}$.
\end{mylem}
\begin{proof}
TODO
% TODO
\end{proof}

\begin{mylem}\label{lemma:minextorder}
% TODO orderset only defined on simple
For any minimal sequence $S\neq\cbrk$, $\orderset{S} = \extset{S}$.
\end{mylem}
\begin{proof}
TODO
% TODO
\end{proof}

\begin{mylem}
All sequences in $\orderset{S}$ of a simple sequence $S$ are equivalent.
\end{mylem}
\begin{proof}
This follows from \cref{simple_is_max,lemma:minextorder}.
\end{proof}

@@

\begin{mycor}\label{workschained}
% An inference rule in the algebra
$\workssign$ can be chained:
if $A_1\in\seqset{A}$, then
$ \worksc{\seqset{A}}{\seqset{B}} \wedge \worksc{\seqset{C}}{A_1} \Rightarrow \worksc{\seqset{A}\cup\seqset{C}}{\seqset{B}}$.
\end{mycor}

@@

\begin{mylem}\label{unrelated_is_independent}
Two different commands that are not assertion commands are independent iff the nodes they change are unrelated:
\[ \cxy(n,V_Y)\indep\czw(m,V_W) \Leftrightarrow n\unrel m \]
if $\cxy(n,V_Y)\neq\czw(m,V_W)$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
We proceed in two steps; first, we prove that
\[ \cxy(n,V_Y)\nindep\czw(m,V_W) \wedge n\unrel m \] would lead to contradiction.
By definition, if the two commands are not independent, then one of the following holds:
\begin{itemize}
\item $\czw(m,V_W)\cc \cqr(o,V_R) \equiv\cbrk$
\item $\cqr(o,V_R)\cc \czw(m,V_W) \equiv\cbrk$
\item $\czw(m,V_W)\cc \cqr(o,V_R) \nequiv \cqr(o,V_R)\cc \czw(m,V_W)$
\end{itemize}
However, from \cref{ax_separate_commute,ax_separate_nobreaks} we also know that
if $n\unrel m$ then none of the above can be true.

@@

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\begin{align*}
\forall\FS: &\cxy(n,C_Y)\indep \czw(m,C_W) \\
&\wedge \cxy(n,C_Y)\FS\neq\fsbroken \\
&\wedge \czw(m,C_W)\FS\neq\fsbroken \\
\Rightarrow &\cxy(n,C_Y)\cc \czw(m,C_W)\FS\neq\fsbroken.
\end{align*}
\end{myax}
\begin{proof}
We name this proposition an axiom for the same reason as above,
but it is easy to provide an indirect proof based on our filesystem model.
Assume that $\cxy(n,C_Y)\cc \czw(m,C_W)\FS=\fsbroken$.
We know $\cxy(m,C_Y)\FS\neq\fsbroken$ so it must be applying 
$\czw(m,C_W)$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if there the input type does not match the filesystem,
but we know $\FS(m)\in\setvx{W}$ and so
$(\cxy(m,C_Y)\FS)(m)\in\setvx{W}$ as based on \cref{unrelated_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxy(m,C_Y)$.
\end{proof}

@@

From \cref{unrelated_is_independent,ax_separate_commute} we know that
the commands in $S$ and $T$ pairwise commute, and so any sequence
that contains the commands from $S$ and $T$ preserving their original order,
but in any permutation between $S$ and $T$ is equivalent to $S\cc T$ on all filesystems.

Let the command that breaks $\FS$ in $T$ be $t$
so that $T=T_0\cc t\cc T_1$.
It is still true that $(T_0 \cc t)\FS\neq\fsbroken$,
and by definition $(S\cc T_0)\FS\neq\fsbroken$,
but $(S\cc T_0\cc t)\FS=\fsbroken$.
Also, from above we know that $S\cc T_0\equiv T_0\cc S$
and so $(T_0 \cc S)\FS\neq\fsbroken$.

@@

We can repeat this step for $s_2$, the next command in $S$,
and from 
$(T_0 \cc s_1\cc t)\FS\neq\fsbroken$
and
$(T_0 \cc s_1\cc s_2)\FS\neq\fsbroken$
arrive at
$(T_0 \cc s_1\cc s_2\cc t)\FS\neq\fsbroken$.
This can be repeated until $S$ is exhausted and we get
$(T_0 \cc S\cc t)\FS\neq\fsbroken$, which is a contradiction.

\end{proof}

From this and \cref{workschained} it follows that:

\begin{mycor}\label{r_indep_combine}
% An inference rule in the algebra
The combination of independent sequences continues to work under the same conditions:
\[ \worksc{S}{\seqset{U}} \wedge \worksc{T}{\seqset{U}} \wedge S\indep T \Rightarrow \worksc{S\cc T}{\seqset{U}} \]
\end{mycor}

@@

\begin{mydef}{Inverse commands and sequences.}
The inverse of command $\cxy(n,V_Y)$ is $\cxy(n,V_Y)^{-1} = \cyx(n,V_X)$
where $V_X$ is an arbitrary value from $\setvx{X}$.
We write $S^{-1}$ for the inverse of sequence $S$, which consists of the inverses of the commands in $S$
in reverse order.
%
Clearly, $(S^{-1})^{-1}\equiv S$ for any sequence.
\end{mydef}

\begin{mylem}\label{r_invmove}
A common prefix of sequences can be moved to the other side of $\workssign$ by inverting it:
\begin{gather*}
\worksc{B\cc \seqset{A}}{\seqset{S}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{S}} \\
\worksc{\seqset{A}}{B\cc \seqset{S}} \Rightarrow \worksc{B^{-1}\cc \seqset{A}}{\seqset{S}}
\end{gather*}
% TODO Proof
TODO.
Sequences are bijections on filesystems over type-equality.
TODO: $A,S,B$ need to be simple?
\end{mylem}

\begin{mylem}\label{indep_prefix_combine}
The combination of sequences with a common head and independent tail continues to work:
\[ \worksc{A\cc B}{\seqset{S}} \wedge \worksc{A\cc C}{\seqset{S}} \wedge B\indep C \Rightarrow \worksc{A\cc B\cc C}{\seqset{S}} \]
\end{mylem}
\begin{proof}
Based on \cref{r_invmove} we know
$\worksc{B}{A^{-1}\cc \seqset{S}}$ and $\worksc{C}{A^{-1}\cc \seqset{S}}$,
and from \cref{r_indep_combine} we know
$\worksc{B\cc C}{A^{-1}\cc \seqset{S}}$. Applying the second line of \cref{r_invmove} we get
$\worksc{(A^{-1})^{-1}\cc B\cc C}{\seqset{S}}$, which proves our lemma as $(A^{-1})^{-1}\equiv A$.
\end{proof}

>>>>

\newcommand{\cbrk}{\mathtt{break}} % 'break' command
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyx}{\fscommand{Y}{X}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\cxw}{\fscommand{X}{W}}
\newcommand{\czw}{\fscommand{Z}{W}}
\newcommand{\cqr}{\fscommand{Q}{R}}
\newcommand{\cqq}{\fscommand{Q}{Q}}

@@

\newcommand{\eqext}{\sqsubseteq} % equivalent to
\newcommand{\eqnrw}{\sqsupseteq} % extended by
\newcommand{\nequiv}{\not\equiv}
\newcommand{\indep}{\mathrel{\wr\wr}} % Independent commands, sequences
\newcommand{\unrel}{\indep} % Unrelated nodes
\newcommand{\nindep}{\,{\centernot{\wr\wr}}\,} % not \indep
\newcommand{\nunrel}{\nindep} % not \unrel

\newcommand{\workssign}{\mathbf{w}}
\newcommand{\works}[1]{\workssign({#1})}
\newcommand{\worksc}[2]{\workssign({#1}\mathrel{|}{#2})}

\newcommand{\typeeq}{\stackrel{t}{=}}
\newcommand{\ntypeeq}{\centernot{\stackrel{t}{=}}}

\newcommand{\emptyseq}{\lambda} % empty sequence

\newcommand{\orderset}[1]{\vec{\wp}({#1})}
\newcommand{\seqset}[1]{\underline{#1}} % Set of sequences

@@

To model this, in $\setv$ we distinguish between files ($\setf$) and directories ($\setd$), that is,
if $\setb = \{\empt\}$ then:
\[ \setv = \setb \cup \setf \cup \setd \]
The tree propety can then be defined as
\begin{align*}
\forall n\in\setn: &\FS(n) \neq \empt \\ % TODO alignment problem
&\quad\wedge \parent(n) \neq \topnode \\
&\Rightarrow \FS(\parent(n)) \in \setd 
\end{align*}
This means that as we move down from the root of a tree of nodes,
the types of values we encounter in the filesystem can only change according to the following
transition diagram:

@@

% <<<<<<<< TODO Rewrite: no algebra
%  about which we will reason
% using our algebra.
% We will aim to draw conclusions or judgements about commands and sequences of commands
% based on the algebra we are aiming to construct.
% For example, if we believe that two sequences of commands are equivalent as their effect
% is the same, then we aim to be able to derive this from the axioms and inference rules of the
% algebra.
% As the algebra operates on commands only,
% these judgements, by definition, need to hold regardless of the filesystem the commands are applied to.
% Accordingly, we expect that the more information one encodes into the commands and the sequences,
% the better predicitons we will be able to make using our algebra,
% as then the commands and sequences will be more specific, and will select a smaller subset
% of potential filesystems on which they can meaningfully operate.
% >>>>>>>>>> to here

...

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(n,V)$, $edit(n,V)$ and $remove(n)$ where $n\in\setn$ and $V\in\setv$ (but $V\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC:alg} we know that a useful set of axioms
will in some cases need to distinguish between edits that result in directories ($edit(n,D)$) and
ones that result in files ($edit(n,F)$), and treat them as separate commands.
Indeed, Bill Zissimopoulos suggested (\cite{BZ}) that extending this distinction to more commands
utlimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($D$, $F$ or $\empt$) in the commands is definitely useful.
At the same time, there is never any need to encode information about the
\emph{exact} output value of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

@@

We therefore propose to have a command for each pair of types.
For want of a better system, we will name our commands by concatenating
two of $\fsregcommandchar{D}$, $\fsregcommandchar{F}$ and $\fsregcommandchar{\empt}$, 
where the left sign denotes the type of value
required in the filesystem before the command is applied, and the right sign notes the type
afterwards. For example, $mkdir(n,D)$ is $\cbd(n,D)$, and $rmdir(n)$ is $\cdb(n)$.

A command can only succeed if the original value at node $n$ has a type that matches
the input type of the command. If this is not the case, or if the resulting
filesystem no longer has the tree property, then we say that the command
{\em breaks} the filesystem. Broken filesystems are considered equal
(but not equal to any working filesystem), and we note them by $\FS=\fsbroken$.

@@

In general, when we describe multiple or unknown commands, we may substitute one or both
of their types with variables, as we did with $\cxy(n,V_Y)$. In this context,
if we write that $\cxy=\czw$, we mean that the input and output types
of the commands are the same ($X=Z$ and $Y=W$), while a full equiality 
($\cxy(n,V_Y)=\czw(n,V_W)$)
implies
that their output value is also the same.

@@

The commands we aim to investigate do not occur in isolation:
for example, when comparing filesystems to determine the commands needed to transform one into the other,
we receive a set of commands with uncertain ordering.
Applying commands to a filesystem, however, happens in time,
so for this we need a sequence of commands.
We use $\cc$ to concatenate commands to form a sequence, or concatenate sequences to form a longer sequence:
$S=\cbd(n)\cc \cbf(m,F)$, or $S'=S\cc\cbb(o)$.
We write $\emptyseq$ for the empty sequence.

TODO: Sequences of commands, in the usual way,  form a free
semigroup

@@

Our aim is to derive information about the effects of sequences
of commands independently of that actual filesystems.
In order to make this possible, we investigate the smallest building
blocks of sequences: pairs of commands that act on a filesystem directly one after the other.
This approach is useful as there are a limited number of command pairs,
because, as we argued above, we can disregard the exact output values of commands apart from their type,
and we can also abstract the relationship between the nodes in the two commands
to a finite number of cases.

@@

Below we list a number of statements derived using this method.
Although they can be proven by investigating all relevant possibilities
in a filesystem, we refer to these as axioms because in a secondary algebraic model
that describes commands without reference to an underlying filesystem,
these would function as meta-rules generating statements we would accept as true.
See section (TODO) for more about such a model.

@@

In this section we describe an algorithm with which all possible orderings of a
simple set of commands can be generated, but before doing so, let us consider the following lemma.

\begin{mylem}\label{connected_changes}
If a simple set of commands, which can be applied to a filesystem in some order without breaking it,
contains commands on $n$ and $\nn$ where $\nn\descendant n$
but $\nn\neq\parent(n)$, then the set must also contain a command on $\parent(n)$.
\end{mylem}
\begin{proof}
We will prove this statement for simple sequences, as if all sequences must contain a command on $\parent(n)$
under the given conditions, then so must sets because otherwise there would be no order in which thay can be
applied to any filesystem.

Let $A$ be a sequence that satisfies the conditions;
we therefore know that it contains a command $\cxy(n,V_Y)$ on $n$
and another command $\czw(\nn,V_W)$ on $\nn$.
We use an indirect proof and assume that there are no commands on $\parent(n)$ in $A$.
Next, we create a new sequence $A'\equiv A$ in which $\cxy$ and $\czw$ are next to each other.
If they are already next to each other in $A$, there is nothing to do.
Otherwise, consider the command next to $\cxy(n,V_Y)$ in the direction where $\czw(\nn,V_W)$ is.
Let this command be $\cqr(m,V_R)$.
If $\czw$ is to the right, then $A$ looks like the following:
\[ A = \cdots\cc\cxy(n,V_Y)\cc\cqr(m,V_R)\cc\cdots\cc\czw(\nn,V_W)\cc\cdots \]
If $n\unrel m$, reverse $\cxy$ and $\cqr$. Based on \cref{ax_separate_commute} we know that the new
sequence is equivalent to $A$.
Otherwise, we know $m\neq\parent(n)$ as there are no commands on $\parent(n)$, and so
from \cref{ax_distantrel_breaks} we get $A\equiv\cbrk$ which contradicts our assumptions.
(Note that $A$ cannot contains assertion commands.)
By repeating this step we can therefore convert $A$ into $A'$ where $\cxy$ and $\czw$ are neighboring commands.
However, then \cref{ax_distantrel_breaks} applies to the sequence and therefore $A\equiv A'\equiv\cbrk$ which
is again a contradiction.
\end{proof}

By applying \cref{connected_changes} multiple times we see that if a simple set of commands
contains two commands on related nodes, then it must contain a command on each
node that is between these two nodes in the node tree.

@@

To order the commands in the component,
select the command with the topmost node, so that there would be no command
on its parent node (if there is one). 
If the command removes content at this node (it is of the form $X\empt$), 
then it must follow all other commands in the component, and
all parent--child pairs in the whole component must be destruction pairs.
Otherwise, the command with the topmost node needs to precede the other commands,
and the component must be made of construction pairs.
These follow from the patterns of construction and destruction pairs, which
at no point can mix without causing any sequence to break all filesystems.

Then move the command with the topmost node to the beginning or the end of
the sequence we are creating out of the component,
and recursively start the whole algorithm again on the commands in the component that remain,
potentially splitting them into sub-components.
The result of this process is a set of possible orders for each component,
which are then combined into a set of possible orders for the whole set of commands.

@@

As a specific case, we can regard a sequence of commands $S$
as a set and determine its $\orderset{S}$.

From the algorithm we also see that if a simple sequence $S$ works, then
all sequences in $\orderset{S}$ are equivalent to it.
As the sequences contain a single command on each node,
their output must be the same on the same input unless
they break the filesystem.
However, the algorithm ensures that that will not be the case.
% TODO
(TODO: Is this enough?)
Therefore:

\begin{mycor}\label{simple_reorder_equiv}
For any simple sequence $S\neq\cbrk$,
all sequences in $\orderset{S}$ are equivalent.
\end{mycor}

@@

\begin{mycor}\label{workschained}
% An inference rule in the algebra
$\workssign$ can be chained:
if $\seqset{A'}\subset\seqset{A}$, then
$ \worksc{\seqset{C}}{\seqset{A'}} \wedge \worksc{\seqset{A}}{\seqset{B}} \Rightarrow \worksc{\seqset{A}\cup\seqset{C}}{\seqset{B}}$.
\end{mycor}

@@

\begin{mylem}\label{unrelated_is_independent}
Two different commands that are not assertion commands are independent iff the nodes they change are unrelated:
\[ \cxy(n,V_Y)\indep\czw(m,V_W) \Leftrightarrow n\unrel m \]
if $\cxy(n,V_Y)\neq\czw(m,V_W)$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
We proceed in two steps; first, we prove that
\[ \cxy(n,V_Y)\nindep\czw(m,V_W) \wedge n\unrel m \] would lead to contradiction.
By definition, if the two commands are not independent, then at least one of the following holds:
\begin{itemize}
\item $\czw(m,V_W)\cc \cqr(o,V_R) \equiv\cbrk$
\item $\cqr(o,V_R)\cc \czw(m,V_W) \equiv\cbrk$
\item $\czw(m,V_W)\cc \cqr(o,V_R) \nequiv \cqr(o,V_R)\cc \czw(m,V_W)$
\end{itemize}
However, from \cref{ax_separate_commute,ax_separate_nobreaks} we also know that
if $n\unrel m$ then none of the above can be true.

@@

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\begin{align*}
\forall\FS: &\cxy(n,C_Y)\indep \czw(m,C_W) \\
&\quad\wedge \cxy(n,C_Y)\FS\neq\fsbroken \\
&\quad\wedge \czw(m,C_W)\FS\neq\fsbroken \\
&\Rightarrow \cxy(n,C_Y)\cc \czw(m,C_W)\FS\neq\fsbroken.
\end{align*}
\end{myax}
\begin{proof}
We name this proposition an axiom for the same reason as above,
but it is easy to provide an indirect proof based on our filesystem model.
Assume that $\cxy(n,C_Y)\cc \czw(m,C_W)\FS=\fsbroken$.
We know $\cxy(m,C_Y)\FS\neq\fsbroken$ so it must be applying 
$\czw(m,C_W)$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if the input type does not match the filesystem,
but we know $\FS(m)\in\setvx{W}$ and so
$(\cxy(m,C_Y)\FS)(m)\in\setvx{W}$ as based on \cref{unrelated_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxy(m,C_Y)$.
\end{proof}

@@

From \cref{unrelated_is_independent,ax_separate_commute} we know that
the commands in $S$ and $T$ pairwise commute, and so any sequence
that contains the commands from $S$ and $T$ and preserve their original partial order
is equivalent to $S\cc T$ on all filesystems.

Let the command that breaks $\FS$ in $T$ when applying $S\cc T$ be $t$
so that $T=T_0\cc t\cc T_1$.
It is still true that $(T_0 \cc t)\FS\neq\fsbroken$,
and by definition $(S\cc T_0)\FS\neq\fsbroken$,
but $(S\cc T_0\cc t)\FS=\fsbroken$.
Also, from above we know that $S\cc T_0\equiv T_0\cc S$
and so $(T_0 \cc S)\FS\neq\fsbroken$.

@@

We can repeat this step for $s_2$, the next command in $S$,
and from 
$(T_0 \cc s_1\cc t)\FS\neq\fsbroken$
and
$(T_0 \cc s_1\cc s_2)\FS\neq\fsbroken$
arrive at
$(T_0 \cc s_1\cc s_2\cc t)\FS\neq\fsbroken$.
This can be repeated until $S$ is exhausted and we get
$(T_0 \cc S\cc t)\FS\neq\fsbroken$, which is a contradiction.
\end{proof}

Separately, we prove the following:

@@

\begin{mydef}{Inverse commands and sequences.}
The inverse of command $\cxy(n,V_Y)$ is $\cxy(n,V_Y)^{-1} = \cyx(n,V_X)$
where $V_X$ is an arbitrary value from $\setvx{X}$.
We write $S^{-1}$ for the inverse of sequence $S$, which consists of the inverses of the commands in $S$
in reverse order.
\end{mydef}

\begin{mydef}{$\typeeq$.}
For two filesystems we write $\FS\typeeq\GS$
iff 
at all nodes the types of their values are the same, or they are both broken.
\end{mydef}

From the definitions we can clearly see that
\begin{mycor}\label{negneg_is_typeeq}
$\forall \FS: (S^{-1})^{-1}\FS\typeeq \FS$.
\end{mycor}

\begin{mylem}\label{r_invmove}
A common prefix of sequences can be moved to the other side of $\workssign$ by inverting it:
\begin{gather*}
\worksc{B\cc \seqset{A}}{\seqset{S}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{S}} \\
\worksc{\seqset{A}}{B\cc \seqset{S}} \Rightarrow \worksc{B^{-1}\cc \seqset{A}}{\seqset{S}}
\end{gather*}
\end{mylem}
\begin{proof}
This is based on the fact that in our model, unless they break a filesystem,
commands change one type of value into another, but never merge types.
That is,
sequences---as functions mapping filesystems to filesystems---are
essentially bijections over type-equality
with the only ``sink'' being $\fsbroken$:
\begin{align*}
\forall S,\FS,\GS: &S\FS\neq\fsbroken \Rightarrow \\
&(S\FS\ntypeeq S\GS \Leftrightarrow \FS\ntypeeq \GS).
\end{align*}

For
\[ \worksc{B\cc \seqset{A}}{\seqset{S}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{S}} \]
the rest of the proof is illustrated by the diagram below
where Dom($x$), the domain of $x$, represents the set of filesystems
the sequence $x$ does not break, or neither of the sequences in the set $x$ breaks.

\input{inverse-works} % TODO arrows to small

We see that the set of filesystems the sequences in $\seqset{A}$ do not break intersect with the range of $B$.
As $B$ is a bijection between its domain and range, we can use $B^{-1}$ to map this intersection back
onto the domain of $B$.
As $\worksc{B\cc \seqset{A}}{\seqset{S}}$ the domain of $\seqset{S}$ must be a subset of this
projected intersection.
If so, then we can use $B$ to map the domain of $\seqset{S}$, which yields the domain of $B^{-1}\cc \seqset{S}$.
As it is also a part of the domain of $A$, we get $\worksc{\seqset{A}}{B^{-1}\cc \seqset{S}}$.

The second part of the lemma can be proven in a similar way.
\end{proof}
